\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-04 response.tex

\documentclass[12pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[margin=20mm]{geometry}
\usepackage{amsmath,color,amsthm,enumerate,amsfonts,mathtools}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{amsfonts}
\usepackage{paralist}
\usepackage{stmaryrd}

\newcommand{\done}{\textcolor{blue}{done}}
\newcommand{\tba}{\textcolor{red}{to be addresssed}}

@@

2. page 1: ``A tree-decomposition $T$ consists of a tree $T$ and a collection
$T$ '' $\longrightarrow$ Delete the first $T$ and define a $T$-decomposition.

\done

Other comments:

1. abstract: “This leads to analogous results for map graphs, string graphs, graph powers, and nearest neighbour graphs.” $\longrightarrow$ Add “certain types of” or be more specific. Same comment for immediately after Theorem 4.

2. page 1: “A tree-decomposition T consists of a tree T and a collection
T ” $\longrightarrow$ Delete the first T and define a T-decomposition.
>>>>>>> f2dea7a16f3c0af4014722c98524e5a9481eb01b

@@

Page 6. Replace 'parent in $T'$ by 'parent in $T_0$'.

\done

Page 6. The term 'hierarchical decomposition' is not defined.

\textcolor{blue}{We have put `hierarchical' in quotes to indicate it is not a defined word.}

### 2022-03-05 k-planar.tex
added comments about how to fix the 1-planar proof;  it is all fixed if we  define kite in the right way

A \defin{kite} in $G$ is the subgraph $K=G[\{v,w,x,y\}]$ induced by the endpoints of a pair of crossing edges $vw,xy\in E(G)$.  It follows from edge-maximality that every kite is isomorphic to the complete graph $K_4$.
The edges $vw$ and $xy$ are called \defin{spars} of $K$.  The cycle $vxwy$ is called the \defin{sail} of $K$.  It follows from edge-maximality that none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of $G$.

\referee{2}{Page 13. Perhaps I am misunderstanding something, but I do not see why
'none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of
$G$.'  For example, see the attached PDF for a picture where $vw$, $xw$, $wy$,
and $yv$ are all crossed by other edges of $G$.\\
\includegraphics[width=60mm]{kite}}
\note{DW}{I agree with the referee.}
Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.
\note{DW}{This is wrong I think} 
Observe that any spar of $K$ is incident on exactly four \defin{kite faces} of $G$, each of which has three edges and two vertices of $G$ on its boundary.

\referee{2}{Page 13. The attached PDF also shows that the definition of 'kite
face' may not be well-defined.  For example, I do not see why there
cannot be some edges of $G$ 'inside' a kite face.  Even if this is not
the case, I think it is more precise to say that a kite face has 'two
and a half edges' and two vertices of $G$ on its boundary' rather than
'three edges and two vertices of $G$ on its boundary'.}
\note{DW}{I agree with the referee.}

@@

\note{DW}{I agree with the referee.}

@@

Our proof of \cref{1-planar} follows quickly from the following technical lemma, which is an extension of the analagous result for plane graphs \cite{DJMMUW20}.
\begin{lem}
	\label{induction} The setup:
	\begin{compactenum}
		\item Let $G$ and $G'$ be defined as above.
		\item Let $T$ be a BFS spanning tree of $G'$ rooted at some vertex $r$.
		\item For each integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(r,v)=j\}$.
		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that
		\begin{compactenum}
			\item No edge of $F$ is crossed by any edge of $G$; and
			\item $V(F)$ can be partitioned into $P_1,\ldots,P_k$, for some $k\in\{1,2,3\}$ such that for each $i\in\{1,\ldots,k\}$,
			\begin{compactenum}
				\item $F[P_i]$ is a path; and
				\item $|V(P_i)\cap L_j| \le 15$ for all integers $j\ge 0$.
			\end{compactenum}
		\end{compactenum}
		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $F$ or the interior of $F$.
	\end{compactenum}
	Then $N$ has an $H$-partition $\mathcal{P}=\{S_x : x\in V(H)\}$ such that:
	\begin{compactenum}
		\item $H$ is planar;
		\item for all integers $j\ge 0$, and all $x\in V(H)$, $|S_x\cap L_j|\le 15$;
		\item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$; and
		\item $H$ has a tree-decomposition in which every bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
	\end{compactenum}
\end{lem}

@@

	The proof finishes in the same way as the proof in \cite{dujmovic.joret.ea:planar}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
	Elements in each of the $H$-partitions of $N_1,\ldots,N_3$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.  The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~14 in \cite{dujmovic.joret.ea:planar}, except that now each node $x$ has a child for each face $F_i$ of $M^+_x$ that contains a vertex of $G$ in its interior.

>>>>

A \defin{kite} in $G$ is the subgraph $K=G[\{v,w,x,y\}]$ induced by the endpoints of a pair of crossing edges $vw,xy\in E(G)$.  It follows from edge-maximality that every kite is isomorphic to the complete graph $K_4$. \note{DW}{We need to define a kite to be the subgraph with vertex-set $\{v,w,x,y\}$ and edge-set $\{vx,xy,vx,vy,wy,wx\}$ where $vx,vy,wy,wx$ are the edges on the `outside' of the kite-faces. Then it really is a simple $K_4$. I think this is what we intended. We should add a figure showing a kite and kite-faces. }
The edges $vw$ and $xy$ are called \defin{spars} of $K$.  The cycle $vxwy$ is called the \defin{sail} of $K$.  It follows from edge-maximality that none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of $G$.

\referee{2}{Page 13. Perhaps I am misunderstanding something, but I do not see why
'none of the edges $vx$, $xw$, $wy$, or $yv$ are crossed by any other edges of
$G$.'  For example, see the attached PDF for a picture where $vw$, $xw$, $wy$,
and $yv$ are all crossed by other edges of $G$.\\
\includegraphics[width=60mm]{kite}}
\note{DW}{We need to explain more about how edge-maximality fixes this problem. In this case we can draw other $xv$, $vy$, $yw$ $wx$ edges close to the central crossing edges.}
\note{DW}{The graph drawing literature defines a 1-planar graph to be \defin{optimal} if it is obtained from a plane quadrangulation by adding the two crossing edges across each face. Optimal 1-planar graphs have the maximum number of edges out of all $n$-vertex 1-planar graphs.	But I am pretty sure there are edge-maximal 1-planar graphs that are not optimal. We should check that our proof is okay on these examples. }
Thus any edge that is a spar of a kite $K$ is not part of a sail of any kite $K'$.

\referee{2}{Page 13. The attached PDF also shows that the definition of 'kite
face' may not be well-defined.  For example, I do not see why there
cannot be some edges of $G$ 'inside' a kite face.  Even if this is not
the case, I think it is more precise to say that a kite face has 'two
and a half edges' and two vertices of $G$ on its boundary' rather than
'three edges and two vertices of $G$ on its boundary'.}
\note{DW}{This should be fixed if we define kite more precisely. A kite-face is bounded by one edge and two half-edges right?}

@@

\note{DW}{If we define kite as above, this is not a problem}

@@

Our proof of \cref{1-planar} follows quickly from the following technical lemma, which is an extension of the analogous result for plane graphs \cite{DJMMUW20}.
\begin{lem}
	\label{induction} The setup:
	\begin{compactenum}
		\item Let $G$ and $G'$ be defined as above.
		\item Let $T$ be a BFS spanning tree of $G'$ rooted at some vertex $r$.
		\item For each integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(r,v)=j\}$.
		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that
		\begin{compactenum}
			\item No edge of $F$ is crossed by any edge of $G$; and
			\item $V(F)$ can be partitioned into $P_1,\ldots,P_k$, for some $k\in\{1,2,3\}$ such that for each $i\in\{1,\ldots,k\}$,
			\begin{compactenum}
				\item $F[P_i]$ is a path; and
				\item $|V(P_i)\cap L_j| \le 15$ for all integers $j\ge 0$.
			\end{compactenum}
		\end{compactenum}
		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $F$ or the interior of $F$.
	\end{compactenum}
	Then $N$ has an $H$-partition $\mathcal{P}=\{S_x : x\in V(H)\}$ such that:
	\begin{compactenum}
		\item $H$ is planar;
		\item for all integers $j\ge 0$, and all $x\in V(H)$, $|S_x\cap L_j|\le 15$;
		\item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$; and
		\item $H$ has a tree-decomposition in which every bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
	\end{compactenum}
\end{lem}

@@

	The proof finishes in the same way as the proof in \cite{DJMMUW20}.  The paths $P_1,\ldots,P_k$, and $S=V(Y^+)$ become elements of the $H$-partition.
	Elements in each of the $H$-partitions of $N_1,\ldots,N_3$ that intersect $P_1,\ldots,P_k$, or $V(\overline{Y}^+-F)$ are discarded and all the resulting sets are combined to obtain an $H$-partition of $G$.  The desired tree decomposition of $G$ is obtained in exactly the same way as in the proof of Lemma~14 in \cite{DJMMUW20}, except that now each node $x$ has a child for each face $F_i$ of $M^+_x$ that contains a vertex of $G$ in its interior.

	\note{DW}{Check whether Lemma~14 is the right lemma}
